---
title: "Zomato Dataset Analysis"
author: "Vinesh Kumar Gande"
date: "2023-02-03"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


**Dataset link:** Zomato Dataset. (2022, December 23). Kaggle. https://www.kaggle.com/datasets/rajeshrampure/zomato-dataset \
Description of Dataset: The Zomato dataset is imported from Kaggle, this has listing of Bengaluru City Restaurants that have Zomato accounts. We have more than 50000 rows and 17 columns in this dataset, making it pretty sizable.\
For the analysis purpose, i have considered 8 columns, they are,\
1.Restaurant Name: This contains the name of each restaurant\
2.Accepting_Online_Order: This has "yes" or "no", stating the information whether the restaurant accepts online orders or not.\
3.Facility_To_Book_Table: This has "yes" or "no", stating the information whether the restaurant accepts table booking or not.\
4.Customer_Rating: This has the information about the rating given by the customers which is the average of the all the voter surveys.\
5.Number_Of_Votes: The number of people voted for a given restaurant.\
6.Restaurant_Location: The location at which the restaurant is located at.\
7.Restaurant_Type: This lists the type of the restaurant like Casual Dining, Cafe, Quick Bites,Delivery.\
8.Cost_For_Two_People: the average cost for two people that costs for a restaurant.\

Preprocessing Steps:\
1.Selected the features/variables required for the analysis and interpreting the business problem.\
2.Renamed the column names.\
3.Checked for the nulls in each column.\
4.Observed few special characters in the Restaurant Name column and removed the special characters using iconv function.\
5.Identified duplicate values in the columns, then, excluded from the dataframe.\
6.Cleaned the Customer Rating column by removing the duplicates, nulls and the special characters.\
7.Typecasted the datatype of the each column to their respective type.\
8.For the remaining column, if there are duplicates and nulls are present, those were removed from the dataframe.\

```{r warning = FALSE, message=FALSE}
#Loading required libraries
library(readr)
library(dplyr)
data<-read_csv("D:/Course Study/IDMP/HW2/zomato.csv", col_names= FALSE,
               show_col_types=FALSE)
#Loading data from the local library
#problems(data)
```

```{r}
#Choosing the required columns needed for representing the business questions
#location and listed_in(city) columns has the same data, so selecting only 
#location column
df_r<-select(data,Restaurant_Name=X3,Accepting_Online_Order=X4,
             Facility_To_Book_Table=X5,Customer_Rating=X6,
             Number_Of_Votes=X7,Restaurant_Location=X9,
             Restaurant_Type=X10,
             Cost_For_Two_People=X13)
#head(df_r,n=10)
df_r<-df_r[-1,]
# df_r<-select(data,Restaurant_Name=name,Accepting_Online_Order=online_order,
#              Facility_To_Book_Table=book_table,Customer_Rating=rate,
#              Number_Of_Votes=votes,Restaurant_Location=location,
#              Restaurant_Type=rest_type,
#              Cost_For_Two_People='approx_cost(for two people)')
lapply(df_r,function(x) { length(which(is.na(x)))})
#checking nulls across the columns
df_c<-df_r #duplicating the dataframe
head(df_c,n=10)
```
#cleaning Restaurant_Name column\
```{r removing the accents from the Restaurant Name column}
df_c$Restaurant_Name <- iconv(df_c$Restaurant_Name, "UTF-8", "ASCII", sub = "")

# Identify duplicate values based on two or more columns
duplicates <- duplicated(df_c[,c("Restaurant_Name")])

# Remove duplicates based on two or more columns
df_unique <- df_c[!duplicates, ]

# Drop the remaining row accordingly based on two or more columns
df_c <- df_unique[!duplicated(df_unique[,c("Restaurant_Name")]), ]
dim(df_c)
lapply(df_c,function(x) { length(which(is.na(x)))}) 
#checking nulls across the columns

```

#Cleaning Customer_Rating column as it has more nulls when compared to other columns

```{r}
df_c$Customer_Rating <- sub("/5", "", df_c$Customer_Rating) 
#replacing the /5 with empty string, observed the dataset through excel and 
#decided to trim the rating from a/b to a
df_c <- df_c[!grepl("rated", df_c$Customer_Rating,ignore.case = TRUE), ] 
#observed there are multiple unwanted rows which mixed up with invalid data that
#is reviews data is mixed, so cleaning the rows containing rated
df_c$Customer_Rating <- gsub("\\s", "", df_c$Customer_Rating)
#observed few column data being repeated due to white spaces in the string
#(like "1.8" and "1.8 " been identified as separate values), so removed white
#spaces
df_c<-transform(
  df_c,Customer_Rating= ifelse(nchar(Customer_Rating)>3 | is.na(Customer_Rating)
                               | Customer_Rating=="NEW"  |
                                 Customer_Rating=="etc"| Customer_Rating=="BTM"|
                                 Customer_Rating=="dal" |
                                 Customer_Rating=="-", -1, Customer_Rating))
#Removing the rows having the data other than rating
# df_c
df_c <- df_c[grepl("[^''!@#$%^&*()]+", df_c$Customer_Rating), ] #Removing the 
#rows having special characters other than rating
df_c$Customer_Rating <- gsub("[\\[\\]]",-1, df_c$Customer_Rating)

#lapply(df_c,function(x) { length(which(is.na(x)))})

df_c$Customer_Rating<-as.numeric(df_c$Customer_Rating)
# str(df_c)
```

#Cleaning the Accepting_Online_Order column
```{r}
library(stringr)

#removing rows which are not matching with string yes or no(As this column 
#requires yes or no, so cleaning apart from boolean values)
df_c<-transform(
  df_c, Accepting_Online_Order= ifelse(tolower(Accepting_Online_Order)=="yes"|
                                         tolower(Accepting_Online_Order)=="no",
                                       str_to_title(
                                         tolower(Accepting_Online_Order)),
                                       "null"))

df_c<- df_c[!(df_c$Accepting_Online_Order == "null"), ]
#To remove values having null as value(as there are only 16 rows out of 56k rows)
df_c<- df_c[!(is.na(df_c$Accepting_Online_Order)), ] 
#to remove null rows from the dataframe

# unique(df_c$Accepting_Online_Order)
#lapply(df_c,function(x) { length(which(is.na(x)))})
df_c$Number_Of_Votes<-as.numeric(df_c$Number_Of_Votes)
```

#Cleaning Restaurant_Location column
```{r }
#Removing Nulls from Restaurant Location column as there are negligible rows 
#that is 4 out of 11k
df_c<- df_c[!(is.na(df_c$Restaurant_Location)), ]
```

#cleaning restaurant type column
```{r cleaning restaurant type column}
#As the restaurant type column data has different categories clubbed together,
#so replacing nulls with "other" category doesn't make any sense, so removing 
#nulls from restaurant type data.
df_c<- df_c[!(is.na(df_c$Restaurant_Type)), ]
lapply(df_c,function(x) { length(which(is.na(x)))}) 
#to check nulls values column wise

```
#data type change for cost for 2 people column
```{r data type change for cost for 2 people column}
df_c$Cost_For_Two_People <- gsub(",","", df_c$Cost_For_Two_People)
#Removing commas from the column values
df_c$Cost_For_Two_People<-as.numeric(df_c$Cost_For_Two_People)
#changing datatype to numeric from char
str(df_c) #Verifying the data type change
```

```{r Cleaning nulls}
df_c<- df_c[!(is.na(df_c$Cost_For_Two_People)), ]
lapply(df_c,function(x) { length(which(is.na(x)))}) #to check nulls values
#column wise to verify if every column is free from nulls

```

#(Visualisation)

```{r Accepting_Online_Order}
library(ggplot2)
ggplot(df_c,aes(x=Accepting_Online_Order))+geom_bar(width = 0.2)+
  ggtitle("Online order acceptancy rate")
```
**Observations:**\
Among 8685 Restaurants, approximately 59% of the restaurants does not accept online orders. However, there is a negligible difference of 6.67% between the restaurants those who accept the online orders. So, there is a lot of scope for Zomato to expand and convert that "No" rate to "Yes".

```{r Facility to Book Table Distribution }
colors <- rev(RColorBrewer::brewer.pal(name="PuOr",n=3))
yn <- table(df_c$Facility_To_Book_Table)

angles <- cumsum(yn) / sum(yn) * 2 * pi
angles <- append(angles, angles[1])
# Plot the pie chart
percentages <- round(100 * yn / sum(yn), 1)
labels <- paste(names(yn), "\n", percentages, "%", sep="")
pie(yn, labels = labels, radius=1, col=colors, 
    main="Facility to Book Table Distribution", cex=0.8, font=13, init.angle=0, 
    clockwise=TRUE, density=NULL)

for (i in 1:length(yn)) {
  x <- 1.5 + cos(angles[i] + angles[i+1] / 2) * 1.7
  y <- 1.5 + sin(angles[i] + angles[i+1] / 2) * 1.7
  label <- paste(names(yn)[i], "\n", round(100 * yn[i] / sum(yn), 1), "%",
                 sep="")
  text(x, y, label, cex=0.8)
}

```
**Observations:**\
Only 9 percent of the restaurants are facilitating to book a table online, whereas majority of the restaurants are not offering to book a table. The Online delivery app Zomato can make a better plan of action to expand their coverage on online bookings. So that the orders at restaurants can be balanced by both online and offline.\

#Top 10 famous Restaurants
```{r fig.width=15}

library(ggplot2)

Rest_sorted<-df_c[order(-df_c$Number_Of_Votes,-df_c$Customer_Rating),]
top10 <- head(Rest_sorted, 10)
ggplot(top10, aes(x = as.factor(Restaurant_Name), y = Customer_Rating)) +
  geom_bar(stat = "summary",width=0.4)+
ggtitle("Top 10 Famous Restaurants in Bangalore")+
  labs(x="Restaurant Name",y="Rating",size=121)+
  geom_text(label=top10$Customer_Rating,size=7,vjust=1,color="white")+
  theme(axis.text.x = element_text(size=10))
```

**Observations:**\
The above bar graph gives information about the top 10 Restaurants in the city, Bangalore based on user rating and ordered by number of votes. On an average, the Restaurant above 4.2 rating is listed in the top 10.

```{r}
#Correlation between Customer Rating and the Cost for Two People
df_cc<-df_c[(df_c$Customer_Rating!=-1),]
ggplot(df_cc,aes(x=Customer_Rating,y=Cost_For_Two_People))+geom_point()+
  geom_smooth(method=lm,color="darkred",fill="blue")+
  ggtitle("Impact of Customer Rating on the Price")
```

**Observations:**\
There is a positive correlation between the Customer rating and the Cost. Most of the data points lie in between 3.6-4.6 rating. Restaurants with low rating are not charging high amount compared to those with high rating.

```{r fig.width=16, fig.height=12}
# Load the required libraries
library(ggplot2)
library(dplyr)
dflr <- df_c[order(df_c$Number_Of_Votes, decreasing = TRUE),]
dflr<-dflr[!(dflr$Customer_Rating == -1),]
# Group the data frame by location and calculate the max of the rating column
df_grouped <- dflr %>% 
  group_by(Restaurant_Location) %>% 
  summarize(max_rating = max(Customer_Rating), 
            top_rated_restaurant = Restaurant_Name[which.max(Customer_Rating)])

# Plot the graph
ggplot(df_grouped, aes(x = Restaurant_Location, y = max_rating, 
                       label = top_rated_restaurant)) + 
  geom_bar(stat = "identity", fill = "deepskyblue3") +
  geom_text(vjust = -5,hjust=3, color = "white",angle=90,size=3) +
  xlab("Location") +
  ylab("Max Rating")+
  ggtitle("Top Most Rated Restaurant for Each Location across Bangalore") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
